SafeNet: A European database of plant-pollinator interactions

### 2024-02-21 Manuscript/Ms.qmd
edits

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions. Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of the interaction ecology of the different taxonomic groups across space [REF] and environmental gradients [@saunders2023]. In addition, environmental gradients has served as useful tool to explore the impacts of climate on phenological change [REF:Byers 2017]. Importantly, macro-ecological analyses of plant-pollinator networks rarely account for differences in sampling effort [REF] or sampling methodology [REF], ignoring important components that can largely influence our view of plant-pollinator networks [REF]. In addition, despite how relevant are interactions to understand plant-pollinator communities, most communities remain under-sampled [REFERENCES] due to XXXXXXXSpecies occurrences?. Thus, large scale datasets of plant-pollinator interactions so researchers can account for differences across studies...

Species occurrences

Threatened species? Red lists?

Discuss actual TRENDS!

4paragraph

>>>>

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions within their ecological context. Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @saunders2023]. Undoubtedly, macro-ecological studies of plant-pollinator networks makes significant contributions to knowledge. However, they often fail to account for differences in sampling effort and sampling methodology, impacting our view and understanding of plant-pollinator interactions [REFS]. In addition, these networks tend to be incomplete snapshots of the reality because insufficient sampling [@olesen2011; @chacoff2012; @jordano2016], highlighting that strong sampling efforts are needed to capture the missing links of these ecological networks. Thus, providing a high number of informative plant-pollinator networks can help researchers to control for a larger amount of ecological variables when exploring macro-ecological patterns. Further, these different networks require from a taxonomic harmonization process that is generally not shared or reusable [REF]. Providing open, informative and reusable data is neccesary to set the foundation of future plant-pollinator research.
